---
output:
  github_document:
    html_preview: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "tools/readme/README-",
  out.width = "100%"
)
```

```{r eval=TRUE, message=FALSE, warning=FALSE, include=FALSE}
devtools::load_all()
```

**Work in progress, breaking changes are possible! Contributions/input welcome**

# mscgeometR

<!-- badges: start -->
  [![check-standard](https://github.com/rywhale/mscgeometR/actions/workflows/check-standard.yaml/badge.svg)](https://github.com/rywhale/mscgeometR/actions/workflows/check-standard.yaml)
  [![LICENSE](https://img.shields.io/badge/License-MIT-blue.svg)](https://opensource.org/licenses/MIT)
  <!-- badges: end -->

R wrapper for querying the Meteorological Service of Canada's (MSC) [GeoMet webservices](https://eccc-msc.github.io/open-data/msc-geomet/web-services_en).

Includes functionality for querying both the [GeoMet Web Coverage Service (WCS)](https://geo.weather.gc.ca/geomet?lang=en&service=WCS&version=2.0.1&request=GetCapabilities)
as well as the [OGC OpenAPI endpoint](https://api.weather.gc.ca/openapi). 

For more information on both services, please see the 
[official documentation](https://eccc-msc.github.io/open-data/msc-geomet/web-services_en).

## Installation
`mscgeometR` is not available on CRAN, you can install it from GitHub using `remotes`

```{r eval=FALSE}
remotes::install_github("rywhale/mscgeometR")
```

## WCS Data
Queries to the WCS endpoint require you to know the coverage ID of the layer
you're after. You can get a full list of these IDs with `geomet_wcs_capabilities()`

Setting the query parameters requires some familiarity with the product
specification (e.g. resolution). Generally, specifying no "TIME" in the query
will lead to the most recent product being returned. 

Once you have the coverage ID, you can query data like so

```{r wcs}
library(terra)

# Downloads to temporary .tif file
rdpa_file <- geomet_wcs_data(
  # Regional Deterministic Precipitation Analysis
  coverage_id = "RDPA.24F_PR",
  # Set query parameters
  query = list(
    "SUBSETTINGCRS" = "EPSG:4326",
    "OUTPUTCRS" = "EPSG:4326",
    # Res
    "RESOLUTION" = "x(0.09)",
    "RESOLUTION" = "y(0.09)",
    # Bounding box
    "SUBSET" = "x(-96.3553, -89.8336)",
    "SUBSET" = "y(47.3633, 52.0529)",
    "TIME" = paste0(Sys.Date() - 1, "T12:00:00Z")
  )
)

rdpa_rast <- rast(rdpa_file)
crs(rdpa_rast) <- crs("EPSG:4326")

plot(rdpa_rast)
```

`geomet_wcs_data()` downloads the query result to a GeoTIFF file in a temporary
directory. Restarting your R session will clear these temporary files. 

## OGC API Data
You can get a full list of the collections available with

```{r api_col}
geomet_api_collections()
```

The `collection_id` values in the above table can then be used to list potential
query parameters for each collection, e.g. for hydrometric stations

```{r api_queryables}
geomet_api_queryables("hydrometric-stations")
```

Note that boolean variables should be quoted, e.g. `FALSE` -> `"false"`

### Adjusted Homogenized Canadian Climate Data (AHCCD)
```{r ahccd}
# View available stations
head(geomet_ahccd_stns())

# Example query
ahccd_dat <- geomet_ahccd_data(
  station_number = "3011120", 
  period = "year"
  )

head(ahccd_dat)
```

### Climate Station Data
```{r climate}
# View available stations
clim_stns <- geomet_clim_stns()

# Lots of columns, here are the first few
head(clim_stns[, 1:5])

# Example query
clim_data <- geomet_clim_means(
  station_number = "1126070",
  start_date = "2016-01-01",
  end_date = "2016-01-02"
)

head(clim_data)

# Normals
clim_normals <- geomet_clim_normals(station_number = "1126070")

head(clim_normals)
```

### Hydrometric Station Data
```{r hydro}
# View available Ontario stations
hydro_stns <- geomet_hydro_stns(
  query = list(
    "PROV_TERR_STATE_LOC" = "ON"
    )
  )

head(hydro_stns)

# Example means query
hydro_dat <- geomet_hydro_means(
  station_number = "04HA001",
  start_date = "2016-01-01",
  end_date = "2016-01-02"
)

head(hydro_dat)

# Real time query
hydro_rt <- geomet_hydro_realtime(station_number = "04HA001")

head(hydro_rt)
```

### Long Term Climate Extremes (LTCE)
```{r ltce}
# View available stations in Ontario
ltce_stns <- geomet_ltce_stns(
  query = list(
    "PROVINCE_CODE" = "ON"
  )
)

head(ltce_stns)

# Example query
ltce_dat <- geomet_ltce_data(
  station_number = "VSON99V",
  param = "temp"
)

# Lots of columns, here are the first few
head(ltce_dat[, 1:5])
```

### Surface Weather Observations (SWOB)
```{r swob}
swob_dat <- geomet_swob_realtime(
  station_number = "1192948",
  start_date = Sys.Date(),
  end_date = Sys.Date(),
  query = list(
    "_is-minutely_obs-value" = "false"
  )
)

# Lots of columns, here are the first few
head(swob_dat[, 1:5])
```

## Contributing
If you see data sets missing or come across any problems, please [submit an issue](https://github.com/rywhale/mscgeometR/issues).

See [here](https://github.com/rywhale/mscgeometR/blob/main/.github/CONTRIBUTING.md) for general contribution guidelines. 
