---
output:
  github_document:
    html_preview: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "tools/readme/README-",
  out.width = "100%"
)
```

```{r eval=TRUE, message=FALSE, warning=FALSE, include=FALSE}
devtools::load_all()
```

**Work in progress, breaking changes possible**

# mscgeometR

R wrapper for querying the Meteorological Service of Canada's (MSC) [GeoMet webservices](https://eccc-msc.github.io/open-data/msc-geomet/web-services_en).

Includes functionality for querying both the [GeoMet Web Coverage Service (WCS)](https://geo.weather.gc.ca/geomet?lang=en&service=WCS&version=2.0.1&request=GetCapabilities)
as well as the [OGC OpenAPI endpoint](https://api.weather.gc.ca/openapi). 

For more information on both services, please see the 
[official documentation](https://eccc-msc.github.io/open-data/msc-geomet/web-services_en).

## Installation
`mscgeometR` is not available on CRAN, you can install it from GitHub using `remotes`

```{r eval=FALSE}
remotes::install_github("rywhale/mscgeometR")
```

## WCS Data
Queries to the WCS endpoint require you to know the coverage ID of the layer
you're after. You can get a full list of these IDs with `geomet_wcs_capabilities()`

Once you have the coverage ID, you can query data like so

```{r wcs}
library(raster)

# Downloads to temporary file
rdpa_file <- geomet_wcs_data(
  # Regional Deterministic Precipitation Analysis
  coverage_id = "RDPA.24F_PR",
  # Set query parameters
  query = list(
    "SUBSETTINGCRS" = "EPSG:4326",
    "OUTPUTCRS" = "EPSG:4326",
    # Res
    "RESOLUTION" = "x(0.09)",
    "RESOLUTION" = "y(0.09)",
    # Bounding box
    "SUBSET" = "x(-96.3553, -89.8336)",
    "SUBSET" = "y(47.3633, 52.0529)",
    "TIME" = paste0(Sys.Date() - 1, "T12:00:00Z")
  )
)

rdpa_rast <- raster(rdpa_file)
raster::crs(rdpa_rast) <- raster::crs("EPSG:4326")

plot(rdpa_rast)
```

Setting the query parameters above requires some familiarity with the product
specification (e.g. resolution). Generally, specifying no "TIME" in the query
will lead to the most recent product being returned. 

## OGC API Data
You can get a full list of the collections available with

```{r api_col}
geomet_api_collections()
```

The `collection_id` values in the above table can then be used to list potential
query parameters for each collection, e.g. for hydrometric stations

```{r api_queryables}
geomet_api_queryables("hydrometric-stations")
```

Note that boolean variables should be quoted, e.g. `FALSE` -> `"false"`

### Adjusted Homogenized Canadian Climate Data (AHCCD)
```{r ahccd}
# View available stations
head(geomet_ahccd_stns())

# Example query
geomet_ahccd_data(station_number = "3011120", period = "year")
```

### Climate Station Data
```{r climate}
# View available stations
geomet_clim_stns()

# Example query
geomet_clim_means(
  station_number = "1126070",
  start_date = "2016-01-01",
  end_date = "2016-01-02"
)

# Normals
geomet_clim_normals(station_number = "1126070")
```

### Hydrometric Station Data
```{r hydro}
# View available Ontario stations
geomet_hydro_stns(
  query = list("PROV_TERR_STATE_LOC" = "ON")
  )

# Example means query
geomet_hydro_means(
  station_number = "04HA001",
  start_date = "2016-01-01",
  end_date = "2016-01-02"
)

# Real time query
geomet_hydro_realtime(station_number = "04HA001")
```

### Long Term Climate Extremes (LTCE)
```{r ltce}
# View available stations
geomet_ltce_stns(
  query = list(
    "PROVINCE_CODE" = "ON"
  )
)

# Example query
geomet_ltce_data(
  station_number = "VSON99V",
  param = "temp"
)
```

### Surface Weather Observations (SWOB)
```{r swob}
geomet_swob_realtime(
  station_number = "1192948",
  start_date = Sys.Date() - 1,
  end_date = Sys.Date(),
  query = list(
    "_is-minutely_obs-value" = "false"
  )
)
```

## Contributing
If you see data sets missing or come across any problems, please submit an issue
